---
title: "Working with Twitter Data in R"
subtitle: "Resul Umit"
institute: ""
date: "February 2021"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
      ratio: "16:9"
    seal: false
---

class: inverse, center, middle

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, crayon.enabled = TRUE)

library(crayon)
library(fansi)
library(dplyr)
library(ggplot2)
library(stargazer)
library(countdown)

```

<style type="text/css">

.hljs-github .hljs {
    background: #e5e5e5;
}

.inline-c, remark-inline-code {
   background: #e5e5e5;
   border-radius: 3px;
   padding: 4px;
   font-family: 'Source Code Pro', 'Lucida Console', Monaco, monospace;
}


.yellow-h{
   background: #ffff88;
}


.out-t, remark-inline-code {
   background: #9fff9f;
   border-radius: 3px;
   padding: 4px;
   
}

.pull-left-c {
    float: left;
    width: 58%;
    
}

.pull-right-c {
    float: right;
    width: 38%;
    
}

.medium {
    font-size: 75%
    
}

.small {
    font-size: 50%
    }

.action {
    background-color: #f2eecb;
  
}


</style>


# Working with Twitter Data in R

<br>

### Resul Umit

### February 2021

.footnote[

[Skip intro &mdash; To the contents slide](#contents-slide). &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; <a href="mailto:resuluy@uio.no?subject=R Markdown workshop">I can teach this workshop at your institution &mdash; Email me</a>.

]

---
## Who am I?

Resul Umit

- post-doctoral researcher at the University of Oslo

- interested in representation, elections, and parliaments
    - [a recent publication](https://doi.org/10.1177%2F1478929920967588):
    Parliamentary communication allowances do not increase electoral turnout or incumbents’ vote share. 
    
    - [a publication based on Twitter data](https://doi.org/10.1080/13572334.2017.1283166): The voices of Eurosceptic members of parliament (MPs) echo disproportionately louder on Twitter.
    
      - 400,000+ tweets from 1,000+ MPs
      - automated classification, using a bag-of-words approach
      - manual validation
      
    - [an app based on Twitter data](https://resulumit.shinyapps.io/likewise/): LikeWise &mdash; an app that faciliatates searching the tweets that you liked.
    
    - more at [resulumit.com](https://resulumit.com/)

---
## The Workshop &mdash; Overview

- Two days, on how to collect, process, and analyse data from Twitter

  - 200+ slides, 40+ exercises, and time for working on a real project

--

- Based on converting a mock manuscript written in Word to R Markdown

  - plus, improving its reproducibility and version-controlling it
  - with a PDF output in mind

--

- Designed for researchers with basic knowledge of R programming language

  - does not cover programming with R
      - e.g., writing functions    
<br>
  - existing ability to work with data in R will be very helpful
      - but not absolutely necessary &mdash; this ability can be developed during and after the workshop as well
        
---
## The Workshop &mdash; Motivation

- Twitter provides attractive opportunities for academic research
    - a simple search for "twitter data" returns 55,000+ results on Google Scholar, by the end of 2020

--

- Research based on Twitter data requires a set of skills 
    - typically, these skills are not part of academic training


---
## The Workshop &mdash; Motivation &mdash; Opportunities

- Popularity of the network
  - about 350 million users, more than half are daily active
  - your subjects are likely Twitter users
    - e.g., for me, members of parliament

--

- Richness of the data
  - about 500 million tweets per day<sup>*</sup>
  - with up to 90 variables for each tweet

--

- Accesibilty of the data
  - most Twitter data are public
    - i.e., relatively few private profiles
  - programmatic access to Twitter is enabled
    - through API (application programming interface)

.footnote[

<sup>*</sup> The user and usage statistics on this slide are compiled by [BusinessOfApps](https://www.businessofapps.com/data/twitter-statistics).

]

---
## The Workshop &mdash; Motivation &mdash; Challenges

- Research based on Twitter data requires certain skills
    - e.g., availability of API is meaningless for researchers who cannot write code yet
    - processing and analysing are no less challenging without programming skills

- The required skills are often uncovered in the academic training of social scientists
    - e.g., methodology courses
    - but help is available elsewhere
      - collaborate with researchers with the skillset
         - sacrificing some control over your research and research agenda
      - acquire the skills through individual effort
        - e.g., workshops such as this one

---
## The Workshop &mdash; Motivation &mdash; Aims

- To provide you with an understanding of what is possible with R

  - we will cover a large breath of issues, not all of it is for long-term memory
      - one reason why the slides are designed for self study as well    
<br>
  - awareness of what is possible, `Google`, and perseverance are all we need

--

- To start you with acquiring and practicing the skills needed 

  - practice with ready-written code
  - start working on a real project

---
name: contents-slide

## The Workshop &mdash; Contents

.pull-left[

[Part 1. Before You Start](#part1)
   - e.g., considering Twitter for research

[Part 2. Getting the Tools Ready](#part2)
   - e.g., downloading course material

[Part 3. Data Collection](#part3)
   - e.g., acquiring a user's tweets
   
]

.pull-right[
   
[Part 4. Data Preperation](#part4)
   - e.g., creating a tidy dataset of tweets
   
[Part 5. Data Anaysis](#part5)
   - e.g., conducting sentiment analysis
   
]

.footnote[

[To the list of references](#reference-slide).

] 

---
## The Workshop &mdash; Organisation

- Sit in groups of two

  - participants learn as much from their partner as from instructors
  - switch partners after every part

- Type, rather than copy and paste, the code that you will find on these slides

  - typing is a part of the learning process

- When you have a question

  - ask your partner
  - google together
  - ask me

---
class: action

## The Workshop &mdash; Organisation &mdash; Slides

Slides with this background colour indicate that your action is required, for

- setting the workshop up
    - e.g., installing R 
    
- completing the exercises
    - e.g., downloading tweets
    - there are 40+ exercises
    - these slides have countdown timers
        - as a guide, not to be followed strictly
    
`r countdown(minutes = 3, seconds = 00, top = 0)`

---
## The Workshop &mdash; Organisation &mdash; Slides

- Codes and texts that go in R console or scripts .inline-c[appear as such &mdash; in a different font, on gray background]
    - long codes and texts will have their own line(s)

```{r, demonstration, eval=FALSE}

ggplot(data = df, mapping = aes(x = h5_median, y = h5_index, color = subfield)) +
       geom_point() +
       facet_wrap(. ~ branch) +
       scale_colour_discrete(name = "Journal Type", breaks = c(0, 1), labels = c("Generalist", "Subfield"))    

```

---
## The Workshop &mdash; Organisation &mdash; Slides

- Codes and texts that go in R Markdown documents .inline-c[appear as such &mdash; in a different font, on gray background]
    - long codes and texts will have their own line(s)

- Results that come out in output files .out-t[appear as such &mdash; in the same font, on green background]
    - except very obvious results, such as figures and tables
    
--
    
- Specific sections are .yellow-h[highlighted yellow as such] for emphasis
    - these could be for anything &mdash; codes and texts in input, results in output, and/or texts on slides
    
--

- The slides are designed for self-study as much as for the workshop
    - *accessible*, in substance and form, to go through on your own

---
name: part1
class: inverse, center, middle

# Part 1. Before You Start

.footnote[

[Back to the contents slide](#contents-slide).

]

---
## Considerations &mdash; Reseach Questions & Hypotheses

- Ideally, we already have one or more research questions, hypotheses
  - developed prior to data collection on Twitter and analysis
    - based on, e.g., theory, claims, observations

--

- Not all questions can be answered with Twitter data
  - see relevant literature for what works, what does not
    - e.g., for political science, the review by [Jungherr (2016)](https://doi.org/10.1080/19331681.2015.1132401)
    - for public health, the review by [Sinnenberg et al. (2017)](https://doi.org/10.2105/AJPH.2016.303512)

---
## Considerations &mdash; Potential Biases

There are two sources of potential bias in Twitter data

- sampling
  - Twitter users are not representative of the people out there
     - see, for example, [Mellon and Prosser (2017)](https://doi.org/10.1177%2F2053168017720008)
  - what users (do not) tweet about has a strategic component [Umit (2017)](https://doi.org/10.1080/13572334.2017.1283166)


- 
mediation
  - the behaviour on Twitter is mediated through written and unwritten rules
    - e.g., the function to like, but no dislike
      - does this systematically bias the replies towards negative?
    - e.g., the common use of the like function as a bookmark
      - what would a study of likes measure?

---
## Considerations &mdash; Constraints over Data Access

- Twitter has restrictions on data access
    - how much data can you download through API
        - how quickly, frequently, how far dating back, *etc*
        
--

- These restrictions vary from one operation to another, a lot!
    - e.g., historical tweets from all users, including a word
        - up to 18,000 tweets
        - posted up to seven to nine days prior to search
    
    - e.g., historical tweets from a given user
        - up to 3,200 most recent tweets
        - posted anytime
        
    - e.g., live tweets, including a word
        - up to 1% of the total volume
    
.footnote[

We will discuss these limitations in more detail in [Part 2](#part2). 

]

---
## Considerations &mdash; Constraints over Data Access

- Twitter has restrictions on data access
    - how much data can you download through API
        - how quickly, frequently, how far dating back, *etc*

- These restrictions vary from one operation to another, a lot!
    - e.g., historical tweets from all users vs. from a given user

- Twitter might change the rules of the game anytime
    - making these restrictions more or less stricter

---
## Considerations &mdash; Constraints over Data Redistribution

- Twitter restricts content redistribution
   - only Tweet IDs can be made publicly available in datasets
      - not tweets themselves
 - see [Twitter Developer terms](https://developer.twitter.com/en/developer-terms/agreement-and-policy) for further details
 
- Reproducibility of research based on Twitter data is limited in practice
    - i.e., reproducibility after publication, by others
    - technically, they can retrieve the tweets with IDs
        - in practice, this is very demanding, and may even be impossible
            - e.g., the tweets, accounts might be deleted

---
## Considerations &mdash; Changes in the Twitter API

- Twitter is currently switching to a new API
    - replacing API v1 with API v2
    - see https://developer.twitter.com/en/docs/twitter-api/early-access

--

- Existing codes to collect tweets may or may not be affected, depending on
   - how the API v2 will look in the end
       - it is still a work in progress
   - whether and how the `rtweet` package<sup>*<sup> will adopt

--

- Not all changes are bad
   - among others, API v2 promises
       - 'to serve the unique needs and challenges of academic researchers'
           - see https://developer.twitter.com/en/solutions/academic-research
       - unlimited access to all public tweets
   
.footnote[

<sup>*<sup> This is the R package that we will use to collect tweets. More details are in [Part 2](#part2). 

]

---
## Considerations &mdash; Law and Ethics

- It may be illegal and/or unethical to collect and analyze twitter data without user consent
    - Twitter has no problem with it, but
    - your country and/or (funding) institution may think otherwise
    
--
    
- Check the rules that apply to your case

- Reflect on whether using Twitter data for research is ethical

---
## Considerations &mdash; Data Storage

Twitter data might require 

- large amounts of digital storage space
    - typically big data
       - many tweets, variables
    - e.g., a dataset of 1M tweets requires about 250MB 
        - when stored in .Rdata format

--

- private, safe storage spaces
    - due to Twitter's terms
    - but also local laws, institutional requirements
    
---
## Considerations &mdash; Language and Context

- Some tools of analysis are developed for a specific language and context
   - e.g., dictionaries for sentiment analysis, 
       - e.g., in English, for political texts
   - these may not be useful, valid for different languages, and context
   
--

- Some tools of analysis are developed for general context
   - e.g., a dictionary for sentiment analysis in general
   - these may not be useful, valid for a specific context
      - e.g., political science






---
name: part2
class: inverse, center, middle

# Part 2. Getting the Tools Ready

.footnote[

[Back to the contents slide](#contents-slide).

]

---
class: action

## Workshop Slides &mdash; Access on Your Browser

- Having the slides on your own machine might be helpful
    - flexibility to go back and forward if you need
    - ability to scroll across long codes on some slides

- Access at <https://resulumit.com/teaching/twtr_workshop.html>
  - produced with the `xaringan` package (Xie, 2020)


---
class: action

## R &mdash; Download from the Internet and Install

- Programming language of this workshop
  - created for data analysis, extending for other purposes
      - e.g., accessing APIs
  - allows for all three steps in one environment
      - collecting, processing, and analysing Twitter data
  - an alternative: [python](https://www.python.org/)
  
- Optional, if you have R already installed
  - consider updating it if it is not up to date
    - type the `R.version.string` command in R to check the version of your copy
    - the latest version is R 4.0.3

- Download R from [https://cloud.r-project.org](https://cloud.r-project.org)
    - choose the version for your operating system

---
class: action

## RStudio &mdash; Download from the Internet and Install

- Optional, but highly recommended
  - facilitates working with Twitter data in R

- A popular integrated development environment (IDE) for R
  - an alternative: [GNU Emacs](https://www.gnu.org/software/emacs/)

- Download RStudio from [https://rstudio.com/products/rstudio/download](https://rstudio.com/products/rstudio/download)
    - choose the free version
    - if you have RStudio already installed on your machine, check if a newer version is available

---
class: action
name: rstudio-project

## RStudio Project &mdash; Create from within RStudio 

- RStudio allows for dividing your work with R into separate projects, each with own history etc.

  - [this page](https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects) has more information on why projects are recommended

<br>

- Create a new RStudio project from the RStudio menu:

> `File -> New Project`

---
class: action

## R Packages &mdash; Install from within RStudio

```{r eval=FALSE, tidy=FALSE}
install.packages(c("rtweet", "tidytext", "dplyr", "ggplot2"))

```

--

- `rmarkdown` (Allaire et al., 2020), for automating the process of converting R Markdown documents into other formats

---
class: action

## R Packages &mdash; Install from within RStudio

```{r eval=FALSE, tidy=FALSE}
install.packages(c("rmarkdown", "tinytex", "dplyr", "stargazer", "ggplot2"))

tinytex::install_tinytex() #<<

```

- `rtweet` (Kearney, 2020), for collecting tweets
    - alternatives: `twitteR` (Gentry, 2015)

- `dplyr` (Wickham, 2020a), for data manipulation
    - alternatives: e.g., `base`, `data.table` (Dowle & Srinivasan, 2019)

--
    
- `ggplot2` (Wickham, 2020b), for figures
    - alternatives: e.g., `base`, `plotly` (Sievert et al., 2020)

---
class: action

## Twitter &mdash; Open an Account

Sign up for GitHub at [https://github.com](https://github.com/)

- registering an account is free

- usernames are public
      - either choose an anonymous username (e.g., `asdf029348`)
      - or choose one carefully &mdash; it becomes a part of users' online presence
      
- usernames can be changed later
        

---
class: action

## R Markdown Cheat Sheet &mdash; Download from the Internet

Downloading process can be initiated from within RStudio

- follow from the RStudio menu

> `Help -> Cheatsheets -> R Markdown Cheat Sheet` 


---
## Other Resources<sup>*</sup>

- Pandoc User's Guide

   - available at [https://pandoc.org/MANUAL.html](https://pandoc.org/MANUAL.html)

- R Markdown: The Definitive Guide (Xie et al., 2019)

   - open access at [https://bookdown.org/yihui/rmarkdown](https://bookdown.org/yihui/rmarkdown/)
    

- R for Data Science (Wickham and Grolemund, 2019)
  
   - open access at [https://r4ds.had.co.nz](https://r4ds.had.co.nz/)
   
.footnote[

<sup>*</sup> During the workshop, R Markdown Cheat Sheet is likely to be more helpful than these resources, which I recommend to be consulted after the workshop.

]

---
name: part3
class: inverse, center, middle

# Part 3. Data Collection

.footnote[

[Back to the contents slide](#contents-slide).

]

---
name: part4
class: inverse, center, middle

# Part 4. Data Preperation

.footnote[

[Back to the contents slide](#contents-slide).

]

---
name: part5
class: inverse, center, middle

# Part 5. Data Analysis

.footnote[

[Back to the contents slide](#contents-slide).

]

---
class: inverse, center, middle

# References

.footnote[

[Back to the contents slide](#contents-slide).

]

---
name: reference-slide
## References



Dowle, M. and Srinivasan, A. (2019). [data.table: Extension of 'data.frame'](https://cran.r-project.org/web/packages/data.table/index.html). R package, version 1.12.8.

Gagolewski, M. (2020). [stringi: Character String Processing Facilities](https://cran.r-project.org/web/packages/stringi/index.html). R package, version 1.4.6.

Jungherr, A. (2016). [Twitter use in election campaigns: A systematic literature review](https://doi.org/10.1080/19331681.2015.1132401). Journal of Information Technology & Politics, 13(1), 72-91.

Mellon, J., & Prosser, C. (2017). [Twitter and Facebook are not representative of the general population: Political attitudes and demographics of British social media users](https://doi.org/10.1177%2F2053168017720008). Research & Politics, July-September 2017, 1--9.

Sinnenberg, L., Buttenheim, A. M., Padrez, K., Mancheno, C., Ungar, L., & Merchant, R. M. (2017). [Twitter as a tool for health research: A systematic review](https://doi.org/10.2105/AJPH.2016.303512). American Journal of Public Health, 107(1), e1-e8.

Umit, R. (2017). [Strategic communication of EU affairs: An analysis of legislative behaviour on Twitter](). Journal of Legislative Studies, 23(1), 93-124.

Wickham, H. and Grolemund, G. (2019). R for data science. O'Reilly. Open access at [https://r4ds.had.co.nz](https://r4ds.had.co.nz/).

Wickham, H., Chang, W., Henry, L., Pedersen, T. L., Takahashi, K., Wilke, C., Woo, K., Yutani, H. and Dunnington, D. (2020a). [dplyr: A grammar of data manipulation](https://cran.r-project.org/web/packages/dplyr/index.html). R package, version 0.8.5.

---
## References

Wickham, H., François, W., Henry L. and Müller, K. (2020b). [ggplot2: Create elegant data visualisations using the grammar of graphics](https://cran.r-project.org/web/packages/ggplot2/index.html). R package, version 0.8.5.






---
class: middle, center

## The workshop ends here.
## Congradulations for making it this far, and
## thank you for joining me!

.footnote[

[Back to the contents slide](#contents-slide).

]
