---
title: 'Data Analysis: Users'
output: html_document
---

```{r setup, include=FALSE}

# setup for this document
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)

# install packages that does not need to be loaded
# install.packages("rmakdown")   # running this document   
# install.packages("textdata")   # sentiment dictionary

# load packages
library(rtweet)        # collecting tweets
library(tidytext)      # working with texts, sentiment analysis
library(doc2concrete)  # concreteness scores
library(dotwhisker)    # visualising regression results
library(tidygraph)     # network analysis
library(ggraph)        # network analysis
library(tidyverse)     # most other operations

# load data
mps <- read.csv("../data/mps.csv", na.strings = "")
twts <- read_rds("../data/tweets.rds")

```

## How to use this document

This R Markdown document includes the exercises for Part 5 of the workshop, [Working with Twitter Data in R](https://resulumit.com/teaching/twtr_workshop.html#1). It introduces each exercise with a subtitle, followed by a code chunk. Type your code into these chunks to complete the exercises. 

If you have the `rmarkdown` package installed, you can see the results in two ways:

- by clicking on the *Knit* button on the RStudio menu
    - this runs the codes in all chunks   
    - the results will appear in a separate HTML document that will pop up
    
- by clicking on the green arrow on the top-right corner of any chunk
    - this runs the code in that chunk only    
    - the result will appear immediately under the relevant chunk

If these exercises are too easy for you, I recommend analysing your own data in parallel. If they are too hard, the comments in the chunks might help you. If they are still too inaccesile for you, you may wish to follow on with the `user_based_answers.Rmd` instead, which include the come with the code.


### Correlates of being on Twitter

Looking at the `mps` dataset, we can see that not all MPs have an account on Twitter. Specifically, 588 out of 650 MPs use Twitter. How can we explain that some of them are not on Twitter?

Use the variables from the `mps` dataset to explore.

```{r, echo=FALSE}

# create a new binary variable with `mutate`, coded as 1 for MPs with an account
mps_temp <- mps %>% 
  mutate(on_twitter = if_else(is.na(screen_name), 0, 1))

# run a regression model
m <- lm(on_twitter ~ female + age + term + government + electorate + majority,
         data = mps_temp)

# plot the regression results with the `dwplot` function
dwplot(m) +
  theme_bw() +
  theme(legend.position = "none") +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_x_continuous(breaks = c(-0.1, 0, 0.1)) 


```

### Who has the most followers?

Recall that the `tweets` dataset includes a variable called `followers_count`, measuring the number of followers that MPs have, at the time of data collection. Based on this variable, visualise the top 20 MPs with the highest number of followers.

Optionally, use the `name` and `party` variables from the `mps` dataset to improve the graph. 

```{r, echo=FALSE}

# reduce the dataset to mp-level, with the distinct() function
twts %>%
  distinct(screen_name, followers_count) %>%

# choose the 20 mps with highest number of followers
  arrange(followers_count, desc = TRUE) %>%
  top_n(20) %>%

# get the full name and party variable from the mps dataset
# join by the screen_name variable, which is common in both datasets
  left_join(., mps, by = "screen_name") %>% 

# plot the data with geom_bar
  ggplot(aes(y = reorder(name, followers_count),
             x = followers_count,
             fill = party)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  ylab("Top 20 MPs\n") + xlab("\nNumber of followers") +
  scale_x_continuous(breaks = c(0, 1000000, 2000000, 3000000),
                     labels = c("0", "1M", "2M", "3M")) +
  scale_fill_manual(name = "",
                    breaks = c("Conservative", "Labour", "Green",
                               "LibDem", "Independent"),
                    values = c("#0087DC", "#DC241f", "#008066", "#FDBB30", "gray"))

```

### Correlates of having more followers

How can we explain that some MPs have more followers than others? Looking at the MP-level correlates of followers might be informative. Use the variables from the `mps` dataset as independent variables, to run a regression model with the number of followers being the dependent variable.

```{r, echo=FALSE}

# reduce the dataset to mp-level, with the distinct() function
mps_temp <- twts %>%
  distinct(screen_name, followers_count) %>%

# get the full name and party variable from the mps dataset
# join by the screen_name variable, which is common in both datasets
  left_join(., mps, by = "screen_name")

# run a regression model
 m <- lm(followers_count ~ female + age + term + government + electorate + majority,
         data = mps_temp)

# plot the regression results with the `dwplot` function
dwplot(m) +
  theme_bw() +
  theme(legend.position = "none") +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_x_continuous(breaks = c(-50000, 0, 50000)) 

```

### Who tweets the most often?

Some are more active than others on Twitter. Visualise the top 20 MPs with the highest number of followers, to see the most active ones.

Optionally, use the `name` and `party` variables from the `mps` dataset to improve the graph. 

```{r, echo=FALSE}

# calculate the total number of tweets by mp
twts %>%
  group_by(screen_name) %>%
  summarise(total_tweets = n()) %>%
  ungroup() %>% 

# choose the 20 mps with highest number of followers
  arrange(total_tweets, desc = TRUE) %>%
  top_n(20) %>%

# get the full name and party variable from the mps dataset
# join by the screen_name variable, which is common in both datasets
  left_join(., mps, by = "screen_name") %>% 

# plot the data with geom_bar
  ggplot(aes(y = reorder(name, total_tweets),
             x = total_tweets,
             fill = party)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  ylab("Top 20 MPs\n") + xlab("\nNumber of tweets, 15-31 Jan 2021") +
  scale_fill_manual(name = "",
                    breaks = c("Conservative", "Labour", "SNP"),
                    values = c("#0087DC", "#DC241f", "#FFFF00"))

```

### Correlates of having more tweets

How can we explain that some MPs tweet more often than others?  Looking at the MP-level correlates of the number of tweets might be informative. Use the variables from the `mps` dataset as independent variables, to run a regression model with the number of tweets posted being the dependent variable.

```{r, echo=FALSE}

mps_temp <- twts %>%
  group_by(screen_name) %>%
  summarise(total_tweets = n()) %>%
  ungroup() %>%

# get the full name and party variable from the mps dataset
# join by the screen_name variable, which is common in both datasets
  left_join(., mps, by = "screen_name")

# run a regression model
m <- lm(total_tweets ~ female + age + term + government + electorate + majority,
        data = mps_temp)

# plot the regression results with the `dwplot` function
dwplot(m) +
  theme_bw() +
  theme(legend.position = "none") +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_x_continuous(breaks = c(-50000, 0, 50000)) 



```

# Who do they talk to?

Conventional wisdom suggests that MPs are more likely to interact with MPs from their own party. Does this hold on Twitter, in terms of who MPs reply to? What is the share of in- and out-party replies on Twitter? 

```{r, echo=FALSE}

# create a temporary, reduced dataset of mps and their parties

mps_temp <- mps %>% 
  select(screen_name, party)

# use the `filter` function to select replies, exclude replies to self
twts %>%
  filter(!is.na(reply_to_screen_name) & screen_name != reply_to_screen_name) %>%
  
# use the `inner_join` function to make sure those replied to are MPs as well
# and to get the party variable for the replied to
  inner_join(., mps_temp, by = c("reply_to_screen_name" = "screen_name")) %>%
  
# use the `left_join` function to get the party variable for the screen_name
  left_join(., mps_temp, by = "screen_name") %>%
  
# code a new binary variable for in-party replies with `mutate`
  mutate(in_party = if_else(party.x == party.y, 1, 0)) %>%
  
# reduce the observations to the top two party
  filter(party.x %in% c("Conservative", "Labour")) %>% 
  
# calculate the group percentages
  group_by(party.x, in_party) %>%
  summarise(n = n()) %>%
  mutate(freq = n * 100 / sum(n)) %>%
  
# plot the shares, in a faceted bar graph
  ggplot(aes(x = factor(in_party), y = freq, fill = party.x,
             alpha = factor(in_party))) +
  geom_bar(stat = "identity") +
  theme_bw() +
  theme(legend.position = "none") +
  facet_wrap(. ~ party.x) +
  scale_x_discrete(breaks = c("0", "1"),
                   labels = c("Out-Party", "In-Party")) +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  scale_fill_manual(values = c("#0087DC", "#DC241f")) +
  scale_alpha_manual(values = c(0.5, 1)) +
  ylab("Replies to\n") + xlab("")

```




```{r, echo=FALSE}



```

```{r, echo=FALSE}



```

```{r, echo=FALSE}



```

```{r, echo=FALSE}



```

```{r, echo=FALSE}



```

```{r, echo=FALSE}



```
